% !TEX root = ../my-thesis.tex
%
\chapter{Conclusion}
\label{sec:conclusion}

This thesis presented an end-to-end pipeline that turns natural-language debate
text into a formally evaluated Dung argumentation framework, from typed
attack extraction through confidence filtering to a Prolog solver under grounded,
preferred, and stable semantics.

Returning to the research questions, RQ1 is answered in the affirmative. The
pipeline runs end-to-end across eight UKP topics and an LLM-generated setting,
and it produces well-defined frameworks whose extensions are computed in a
fraction of a second per topic thanks to the grounded-first reduction, since
the undecided cores never exceeded nine arguments in the evaluation.

RQ2 is answered by the ablation. Typed and binary prompting extract
substantially different attack sets on identical atomized arguments, with
directed Jaccard similarities between 0.08 and 0.71, so the categorization
changes what the model asserts rather than rephrasing it. The structural effect
on the frameworks, however, runs entirely through rebuttal symmetrization.
Without the symmetrized edges, typed and binary frameworks alike collapse to a
single extension on every topic. Typing also changes the error profile, since
on the affected topics it suppresses most of the implausible same-stance
attacks that the binary prompt accepts, while introducing failure modes of its
own. The evaluation against the manual reference annotation confirms that this
filtering pays off in accuracy.

RQ3 is answered by the reference evaluation on the frozen Nuclear Energy
atom set. The typed arm recovers twelve of the sixteen manually annotated
directed attacks at a precision of 0.522, a directed F1 of 0.615 against 0.458
for the binary arm, with a spread of 0.579 to 0.615 across three repeated runs
of the same configuration. The errors are systematic rather than random. The
false positives assert attacks across sub-topic boundaries that the strict
object reading of the annotation guidelines excludes, and the missed attacks
concentrate on hedged contradictions. Typed extraction is thus measurably
closer to the manual annotation than the binary baseline.

RQ4 is answered by the cross-topic analysis. Five of eight topics admit
several preferred and stable extensions, and the formal layer makes explicit
several internally consistent positions that the raw argument set leaves
implicit. The advantage of the enumeration over the
grounded extension varies across topics, since on three topics the grounded
extension already decides every argument.

Overall the results demonstrate the feasibility of integrating modern language
models with established formal argumentation, and they suggest that formal
semantics can provide a transparent reasoning layer on top of LLM-generated
argumentative content, provided that the modelling decisions which create the
formal structure, above all the treatment of rebuttals, are made explicit and
kept under experimental control.

\section{Future Work}
\label{sec:conclusion:future}

Several directions remain open. The most immediate one is to broaden the
empirical basis. The extraction quality should be evaluated against annotated
datasets beyond the single-topic manual annotation of this thesis, ideally
with a second annotator so that inter-annotator agreement can be reported, and
the experiments should be repeated across several models to separate
pipeline-specific from model-specific behaviour.

A second direction deepens the symbolic reasoning side. The typed attacks and confidence
scores, which the current pipeline retains only as metadata outside the formal
semantics, could be consumed directly by a richer formalism such as ASPIC+ or
a weighted bipolar framework, and the symmetrization decision could be made
confidence-dependent instead of unconditional.

A third direction concerns the pipeline in use. The exhaustive pairwise
querying bounds it to small argument sets, so scaling to realistic debates
with hundreds of arguments depends on the cheaper candidate prescreening
sketched in Section~\ref{sec:discussion:tradeoffs}, whose price in missed
conflicts would itself have to be measured. The pipeline should also be
compared qualitatively against a standalone LLM that is asked to summarise the
coherent positions of a debate, which would test whether the formal extensions
capture positions that free generation misses.
