% !TEX root = ../my-thesis.tex
%
\chapter{Method: A Pipeline from Debate Text to Formal Extensions}
\label{sec:method}

This chapter describes the pipeline that turns raw debate text into a formally
evaluated Dung framework. Section~\ref{sec:method:overview} gives an overview,
and the remaining sections describe the five steps in turn, each with the design
decision behind it.

\section{Overview}
\label{sec:method:overview}

The pipeline runs in five steps, summarised in
Figure~\ref{fig:method:pipeline}. Step one acquires arguments and atomizes
them into self-contained claims, step two proposes typed attack candidates for
every pair of claims, step three routes the candidates by confidence and
verifies the borderline cases, step four converts the accepted candidates into
binary Dung attack facts, and step five solves the resulting framework under
grounded, preferred, and stable semantics with a Prolog solver.

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[
    node distance = 6mm and 10mm,
    box/.style = {draw, rounded corners=3pt, fill=blue!8,
                  minimum width=82mm, minimum height=8mm,
                  align=center, font=\small},
    arr/.style = {-Stealth, thick},
  ]
    \node[box]                  (atom)   {Step 1: Argument Acquisition and Atomization\\
                                          \textit{LLM generation or corpus input, split into atomic claims}};
    \node[box, below=of atom]   (batch)  {Step 2: Attack Detection and Classification\\
                                          \textit{propose typed attack candidates per pair}};
    \node[box, below=of batch]  (filter) {Step 3: Confidence Filtering and Verification\\
                                          \textit{route by confidence score}};
    \node[box, below=of filter] (aaf)    {Step 4: AF Conversion\\
                                          \textit{convert accepted attacks to binary Dung facts}};
    \node[box, below=of aaf]    (prolog) {Step 5: Solving under Dung Semantics\\
                                          \textit{grounded / preferred / stable extensions}};
    \draw[arr] (atom)   -- (batch);
    \draw[arr] (batch)  -- (filter);
    \draw[arr] (filter) -- (aaf);
    \draw[arr] (aaf)    -- (prolog);
  \end{tikzpicture}
  \caption{Overview of the pipeline, from raw debate text to formally evaluated
  Dung extensions in five steps.}
  \label{fig:method:pipeline}
\end{figure}

\section{Step 1: Argument Acquisition and Atomization}
\label{sec:method:atomize}

Arguments enter the pipeline in one of three ways. They are generated from
scratch by prompting an LLM to produce $n$ pro and $n$ con arguments on a topic,
extracted from an existing text corpus, or loaded directly from an argument
database. In all cases an argument is represented by a tuple of an identifier, a
stance label, and the argument text.

Natural-language arguments frequently bundle several claims. A compound sentence
such as ``Although nuclear energy produces waste, it is cleaner than coal''
mixes a concession with a main claim, and treating it as one unit would distort
the attack relations extracted next. The pipeline therefore prompts an LLM to
split each argument into self-contained atomic claims under three rules.

\begin{enumerate}
  \item \textbf{Drop concessive fragments.} Every component introduced by
    ``although'', ``while'', ``despite'', and similar markers is a rhetorical
    hedge rather than a standalone argument and is dropped.
  \item \textbf{Resolve pronouns.} A pronoun that clearly refers to a specific
    subject is replaced by that subject. A fragment whose referent cannot be
    recovered is removed, so that every atom is understandable in isolation.
  \item \textbf{Filter non-argumentative content.} A statement that neither
    supports nor opposes the claim under debate is dropped.
\end{enumerate}

A fourth rule keeps the stance intact. Reported or conceded positions of the
opposing side, such as ``opponents argue X'', are not promoted to their own
atom. Read in isolation, such a fragment states a claim of the opposing side
while carrying the stance label of its speaker, so the extraction step would
treat it as a claim the speaker actually holds. Every attack decision that
involves such an atom would then rest on a wrongly attributed claim, for
instance a con argument attacking a supposed pro atom that no pro speaker
asserts.

A concrete case from the evaluation illustrates the step. The contra sentence
on meltdown statistics states that in two out of three meltdowns there would
be no deaths, in one out of five there would be over a thousand deaths, and in
one out of one hundred thousand there would be fifty thousand deaths, and thus
bundles three separable claims. Atomization splits it into three atoms, each
carrying one scenario, so that the extraction step can judge each claim on its
own. Each argument is atomized independently, which allows
parallelization. The resulting set of atomic arguments is denoted
$\mathcal{A}'$ and forms the input to the next step. Every atom keeps the
identifier of its source argument. The first atom keeps the bare identifier
and every further atom appends a letter suffix, so that for instance
\texttt{pro5} can yield the atoms \texttt{pro5} and \texttt{pro5b}, which share
the base \texttt{pro5}. The first atom deliberately does not receive a suffix
of its own, so that an argument that needs no splitting passes through the
step with its identifier unchanged. The full atomization prompt is given in
Appendix~\ref{sec:appendix:prompts}.

\section{Step 2: Attack Detection and Classification}
\label{sec:method:extraction}

With the atomic arguments in place, the pipeline examines every pair in
$\mathcal{A}'$ for a potential attack, regardless of stance, since same-stance
arguments can also conflict. Two pro-nuclear arguments may, for example,
prioritise incompatible safety strategies. Pairs of atoms that share a base are
skipped, since they originate from the same source argument.

Rather than a binary decision, the pipeline prompts the LLM to detect and
categorize each potential attack at once and to return a confidence score and a
justification. Asking for the type rather than a mere yes or no forces the model
to engage with the argumentative structure of the pair instead of pattern
matching on opposite stances, and the score and justification make each decision
traceable and provide a routing signal for the next step. Each candidate carries
four fields.

\begin{lstlisting}[basicstyle=\ttfamily\footnotesize]
direction:   A_ATTACKS_B | B_ATTACKS_A | NONE
attack_type: rebuttal | undercutting | undermining | NONE
confidence:  integer in [0, 100]
reason:      one-sentence justification
\end{lstlisting}

The prompt provides the full definitions of the three attack types from
Section~\ref{sec:foundations:typology} and enforces that if
\texttt{direction} is \texttt{NONE} then \texttt{attack\_type} is \texttt{NONE}
and \texttt{confidence} is zero, which keeps the output internally consistent.
Rebuttals are treated as symmetric, so the reverse edge is added automatically,
as motivated in Section~\ref{sec:foundations:typology}.

Beyond the type definitions, the prompt encodes two rules that emerged from
inspecting misclassifications in early runs, and both target the rebuttal
label, the only label whose misuse creates spurious cycles through
symmetrization. The first rule demands a genuine contradiction. Two claims
rebut each other only if they negate the same proposition, so that one can be
true only if the other is false. Merely taking the opposing stance, addressing
a different sub-topic, or arguing for the other side is explicitly ruled out as
a rebuttal. The second rule blocks a recurring category error. A descriptive or
empirical claim, such as a statistic or a report of what people believe, and a
normative claim about what ought to be done are not logically incompatible and
therefore cannot rebut each other, although the descriptive claim may still
undermine a premise of the normative one. Both rules restrict when the mutual,
cycle-creating attack pattern may be asserted, and every prompt is versioned so
that results are always tied to the exact wording that produced them. A final
revision of the prompt, used by the reference evaluation run of
Section~\ref{sec:evaluation:gold}, adds a third rule that applies to all three
attack types. It rejects a pair whenever both specific claims can be true at
the same time without tension, and it requires an undermining attack to name
the premise it hits. The full
prompts are listed in Appendix~\ref{sec:appendix:prompts}.

The remaining
pairs are grouped into batches of size $b$, which reduces the number of API
calls by a factor of $b$. Pairs for which the
model returns malformed output are retried individually.
Algorithm~\ref{alg:method:extraction} summarises this classification step.

This typed prompt is exactly the component whose value the ablation in
Section~\ref{sec:evaluation:ablation} measures, by comparing it against a binary
variant that omits the attack type, the type definitions, and both rules above,
and asks only whether an attack exists.

\section{Step 3: Confidence Filtering and Verification}
\label{sec:method:filter}

Not all candidates are equally reliable. The confidence score, which is an
indication of the model's certainty rather than a calibrated probability, routes
each candidate through a three-way decision with two thresholds
$\theta_{\text{acc}}$ and $\theta_{\text{ver}}$.

\begin{equation*}
\text{decision}(c) =
\begin{cases}
  \textbf{accept} & c \geq \theta_{\text{acc}} \\
  \textbf{verify} & \theta_{\text{ver}} \leq c < \theta_{\text{acc}} \\
  \textbf{reject} & c < \theta_{\text{ver}}
\end{cases}
\end{equation*}

Borderline candidates in the verify region are re-examined by a dedicated
verifier prompt that asks the model to confirm or reject the attack given the two
arguments and the proposed type, but without access to the original confidence
score or justification. Withholding both keeps the verifier from simply
adopting the extractor's judgment instead of forming an independent one.
Accepted and confirmed candidates pass to the next step,
and rejected ones are discarded.

The thresholds are set to $\theta_{\text{acc}} = 70$ and
$\theta_{\text{ver}} = 50$ by default. They are design choices made after inspecting the
confidence values of early runs, in which the model concentrated its scores at
round values and separated clear attacks, which typically received 80 or more,
from speculative ones in the range around 50 to 65. The thresholds were
deliberately not tuned against any evaluation data, since no annotated labels
existed at design time and tuning them on the data used for the evaluation
would leak information. The verification step exists precisely so that the
acceptance threshold can stay high without silently losing every borderline
case, because the middle band gets a second, independent judgment instead of a
hard cut.

\begin{algorithm}[tb]
\caption{Typed attack candidate extraction}
\label{alg:method:extraction}
\begin{algorithmic}[1]
\Require atomized arguments $\mathcal{A}'$, topic $T$, batch size $b$
\Ensure set of typed attack candidates $C$
\State $\mathit{pairs} \gets \{(a_i, a_j) \mid i < j,\;
  \mathrm{base}(a_i) \neq \mathrm{base}(a_j)\}$
  \Comment{exclude same-base pairs}
\State $C \gets \emptyset$
\For{each $\mathit{batch} \in \mathrm{chunk}(\mathit{pairs}, b)$
  \textbf{in parallel}}
  \State $\mathit{raw} \gets \mathrm{ClassifyTyped}(\mathit{batch}, T)$
    \Comment{direction, type, confidence, reason per pair}
  \If{$\mathit{raw}$ parses as a JSON list}
    \State $C \gets C \cup \mathit{raw}$
  \Else
    \State retry the batch once, then classify missing pairs individually
  \EndIf
\EndFor
\State \Return $C$
\end{algorithmic}
\end{algorithm}

\section{Step 4: AF Conversion}
\label{sec:method:conversion}

Every accepted candidate is converted to a binary Dung fact
$\texttt{att}(A, B)$, deliberately dropping the attack type from the formal
semantics. Unlike gradual QBAF semantics, which assign each argument a numerical
strength, Dung semantics ask which arguments can be jointly accepted, a set-based
rather than numerical question, so binary attack facts are the natural
representation. The attack type, the confidence score, and the justification are
not lost. They are retained as comments and metadata and remain available for
analysis and error inspection.

\section{Step 5: Solving under Dung Semantics}
\label{sec:method:solver}

The framework is solved by an AF solver implemented in Prolog under the three
semantics of Section~\ref{sec:foundations:aaf}, using standard fixed-point and
enumeration procedures~\cite{cerutti2017}. The grounded extension is computed by
iterating the characteristic function from the empty set in polynomial time.
Computing preferred and stable extensions requires enumerating conflict-free
sets, which is exponential in the worst case.

To keep this tractable the solver applies a \emph{grounded-first reduction}.
The underlying idea is not new. Computing the cheap grounded labelling first
and restricting the expensive search to the arguments it leaves undecided is a
standard simplification in computational argumentation~\cite{baroni2011,
charwat2015}, and this section describes its concrete form in the pipeline.
The reduction rests on the inclusion hierarchy from
Section~\ref{sec:foundations:aaf}. Let $G$ be the grounded extension and let
$U$ be the set of arguments that are neither in $G$ nor attacked by $G$. This
set is called the \emph{undecided core} in the following. The name comes from
the labelling view of argumentation semantics~\cite{baroni2011}, in which the
grounded labelling marks the members of $G$ as \textsc{in}, the arguments
attacked by $G$ as \textsc{out}, and precisely the arguments in $U$ as
\textsc{undec}, that is, undecided. Every complete extension, and hence every
preferred and stable extension, contains $G$ and excludes all arguments $G$
attacks, so the only freedom left lies in $U$. Moreover, no argument in $U$
attacks or is attacked by $G$, since an attacker of a member of $G$ is itself
attacked by $G$ by definition of defense, and any argument attacked by $G$ is
excluded from $U$ by construction. Attacks into $U$ from arguments that $G$
attacks are neutralised in every complete extension, because $G$ already
counterattacks the attacker. Proposition~\ref{prop:method:reduction} states the
correctness of the reduction in the form the solver uses. The property follows
from known results on complete labellings~\cite{baroni2011}. A self-contained
proof sketch is given in Appendix~\ref{sec:appendix:proof}.

\begin{proposition}[Grounded-first reduction]
\label{prop:method:reduction}
Let $\sigma$ be the complete, the preferred, or the stable semantics. The
$\sigma$-extensions of $AF$ are exactly the sets $G \cup E$ where $E$ is a
$\sigma$-extension of the framework restricted to $U$, with attacks from
outside $U$ into $U$ discounted as neutralised.
\end{proposition}

The exponential enumeration therefore runs only over $U$ instead of over all
arguments. In the running example of Figure~\ref{fig:foundations:example}, the
grounded extension is $\{d, e\}$, the argument $c$ is attacked by it, and the
enumeration only has to decide the two-element core $U = \{a, b\}$. The effect
on real frameworks is drastic, because the extracted graphs are sparse and
their undecided cores are small. On a framework with 24 atoms from an early
run, a naive enumeration over all arguments exceeded a 40 second timeout, while
the reduction solved the identical framework in 0.32 seconds. Since the
enumeration is exponential in $|U|$ and not in the total number of arguments,
the solver additionally guards the computation with a bound of
$|U| \leq \mathit{MaxU} = 18$, determined empirically from solve times, and
falls back to reporting only the grounded extension if the bound is exceeded or
a timeout is reached. In the evaluation of Chapter~\ref{sec:evaluation}, the
undecided cores stay at nine or fewer arguments and the guard never triggers.
Algorithm~\ref{alg:method:solver} in Appendix~\ref{sec:appendix:proof}
summarises the procedure. The solver is
parametrised by the framework file and returns its results as JSON, which the
Python driver reads back.

\section{Implementation and Logging}
\label{sec:method:logging}

The pipeline is implemented in Python, with the solver in
SWI-Prolog~\cite{swiprolog2012}. Every run writes a self-contained log
directory that makes each reported number traceable to its origin, from the
full run configuration and the version identifier of every prompt down to one
record per queried argument pair and per LLM call.
Appendix~\ref{sec:appendix:repro} lists the contents in detail.

Two details of this infrastructure matter for the experiments. Prompt
versioning ties every result to the exact prompt wording and makes the effect
of prompt revisions measurable instead of anecdotal, while the frozen and
reusable atom sets are what allow the ablation in
Section~\ref{sec:evaluation:ablation} to feed the identical atomized arguments
to both extraction variants and the reference annotation of
Section~\ref{sec:evaluation:gold} to be created on exactly the pair set
the pipeline judged.
