% !TEX root = ../my-thesis.tex
%
\chapter{Related Work}
\label{sec:related}

This chapter reviews work that combines language models with formal or
relational argumentation and positions the present thesis against it. It first
discusses end-to-end pipelines that turn text into argumentation frameworks,
then work that uses LLMs specifically for relation extraction, the use of the
UKP corpus, and finally related formalisms. In short, the thesis adopts the
end-to-end perspective of the pipeline systems below and supplies the
empirical evaluation that the closest prior systems leave undone.

\section{LLM-based End-to-End Argumentation Pipelines}
\label{sec:related:pipelines}

The closest work to this thesis is ARGUS~\cite{argus}, which combines LLM-based
argument mining with the construction of argumentation frameworks and symbolic
solvers in an end-to-end architecture. ARGUS shares the central idea pursued
here, namely to let an LLM handle the noisy natural-language part of the problem
and to delegate the acceptability question to a formal solver. It sets out the
architecture and the Dung-semantics perspective but does not report an empirical
evaluation of the resulting frameworks, so it leaves open how many attacks such a
pipeline accepts, whether the resulting frameworks are cyclic, and how many
extensions they admit. Chapter~\ref{sec:evaluation} answers exactly these
questions.

ArgLLMs~\cite{argLLMs} also augments LLMs with formal argumentative reasoning but
takes a quantitative route. Instead of asking which arguments can be jointly
accepted, it builds Quantitative Bipolar Argumentation Frameworks (QBAFs) in
which each argument carries a numerical strength that gradual semantics propagate
along support and attack edges. The system is evaluated on binary claim
verification, so its output is a single scalar per claim rather than a set of
jointly admissible arguments, and its evaluation targets a downstream task
rather than the structure of the argumentation itself.

A further recent example of an end-to-end LLM system for argument mining is the
work of Pirozelli et al.~\cite{pirozelli2026llm}, which applies LLMs to the
argument mining subtasks within a single system rather than through separate
task-specific classifiers. It reflects the broader trend that LLMs are becoming
the default engine for extraction, but like most argument mining work it stops at
the extracted structure and does not connect the result to Dung semantics or
evaluate the collective acceptability of the mined arguments. The pipeline of
this thesis differs from ArgLLMs and Pirozelli et al. by computing set-based
Dung extensions rather than numerical strengths or raw extracted relations,
and from the Dung-based ARGUS by extracting typed, confidence-filtered attacks
and evaluating the resulting frameworks empirically.

\section{LLMs for Attack Relation Extraction}
\label{sec:related:extraction}

The extraction step of the pipeline, which classifies the relation between every
pair of atomic arguments, is most directly related to work that studies LLMs on
relation-based argument mining. Gorur et al.~\cite{gorur2025relMining} ask
precisely whether large language models can identify the relations, such as
support and attack, that hold between arguments, the subtask on which the present
pipeline depends. They find that suitably prompted LLMs are competitive with and
often better than task-specific relation classifiers, which supports the design
decision of this thesis to let an LLM propose the attack candidates. Their
perspective remains diagnostic, in that they measure how well
LLMs recover these relations in isolation, whereas the present thesis treats
relation extraction as one stage of a larger pipeline and feeds its output into a
formal solver. Two design choices set the extraction here apart. The LLM is asked not only
whether an attack exists but also which of the three attack types of
Section~\ref{sec:foundations:typology} it instantiates, which forces engagement
with the argumentative structure rather than a surface judgment on opposing
stances. Each decision moreover carries a confidence score that routes
borderline cases to a separate verification prompt instead of accepting them
outright. The value of
the typed prompt over a plain binary one is measured directly in the ablation of
Section~\ref{sec:evaluation:ablation}.

\section{Argument Mining on the UKP Corpus}
\label{sec:related:ukp}

The evaluation uses the UKP Sentential Argument Mining
Corpus~\cite{stab2018}, which provides sentence-level stance annotations for
eight controversial topics. Prior work of the supervising group applies
knowledge-based graph argument mining with topic modeling to the same
corpus~\cite{kramer2021}. Their task differs from the one studied here. They
classify \emph{single sentences} as argument or non-argument for a given topic,
without stance and without any relation between arguments, and report an
accuracy of 85 percent alongside an F1 score of 67 percent. Two lessons carry
over to this thesis. The 18 point gap between their accuracy and their F1
illustrates how strongly a majority of negative instances inflates accuracy
on this corpus, which is why the reference evaluation in
Section~\ref{sec:evaluation:gold} reports precision, recall, and F1 on the
attack class rather than accuracy. Their F1 also gives a rough sense of how
demanding the corpus is for a related sentence-level subtask, though it is not
a directly comparable benchmark, since pairwise attack detection is a different
and arguably harder problem with a lower positive rate.

\section{Related Formalisms}
\label{sec:related:formalisms}

Most closely related on the symbolic reasoning side is the recent work of
Alfano et al.~\cite{alfano2026}, which connects LLM-based argument mining with
argumentation and description logics, a family of decidable knowledge
representation logics designed for ontological reasoning. Their system extracts
a Fuzzy Argumentative Knowledge Base (FAKB) from raw text and combines it with
a QBAF. This enriches the formal layer and makes it more expressive but also
heavier, since it
commits to a strength ordering that has to be justified. The present thesis deliberately stays with
Dung's abstract framework instead, since the question of interest is which
arguments can be jointly accepted rather than how strong each argument is, and
without a preference ordering binary attack facts are the natural
representation for it,
as Chapter~\ref{sec:method} argues in more detail. The typed attacks and
confidence scores are not discarded but retained as metadata, which keeps the
door open to a richer formalism such as ASPIC+~\cite{modgil2014} or a QBAF as
future work.

