% !TEX root = ../my-thesis.tex
%
\chapter{Introduction}
\label{sec:intro}

\section{Motivation}
\label{sec:intro:motivation}

Human argumentation is hard to capture with classical, deductive logic.
Instead of deriving truth from fixed premises, a debate proceeds
through arguments and counter-arguments, where a claim can be accepted or
defeated depending on which other claims stand against it rather than being
simply true or false. The same difficulty reappears wherever an automated
system must decide which of several conflicting claims can be held together,
most visibly in language models, whose fluent output can assert a conclusion
that the reasons given do not support.

A concrete debate makes the difficulty tangible. Asked about nuclear energy,
people argue that nuclear power is nearly carbon-free, that meltdowns are
statistically rare, that waste storage remains an unresolved safety problem,
that plants are potential terrorism targets, and that subsidies for nuclear
power crowd out investment in renewables. These claims attack
each other in specific and directed ways, and typically more than one
internally coherent position survives, such as a pro camp built around climate
benefits and a contra camp built around catastrophic risk. Which claims can be
held together consistently is a structural question about the debate, and
answering it requires making the attack relationships explicit.

One established way to formalise arguments is the logic of
argumentation~\cite{krause1995logic}, which provides a language for representing
the structure of reasons that support or undercut a line of reasoning. What
such formalisms leave open is where the arguments and their relations come
from in practice, since the examples in the literature often appear
constructed and providing the arguments frequently seems harder than the
reasoning over them.

Large language models (LLMs) approach the problem from the opposite side. Where
formal argumentation offers structure without content, LLMs produce content
without structure. They can generate reasons for or against a cause in fluent
natural language, but they lack the formal scaffolding that would allow any
systematic analysis of how those reasons relate. A standalone LLM can be asked
to summarise the positions in a debate, yet its answer is another ungrounded
generation without a guarantee of internal consistency. The proposal of this
thesis is to combine the two sides, using the LLM for what it is good at,
namely producing arguments and identifying their attacks, and handing the
resulting structure to a formal layer that is good at reasoning about
consistency. The formal layer makes explicit which sets of arguments can be jointly
accepted, a property that the raw LLM output leaves implicit.

The natural target of such a pipeline is argument sets produced by an LLM or
by any other source, since the pipeline itself is agnostic to where the
arguments come from. The evaluation
nonetheless draws its main argument sets from an annotated corpus rather than
from generation. The reason is methodological. If the same model produced the
arguments and also made the attacks between them explicit, the input would
itself be an artefact of the model, and there would be no independent
reference against which the extracted structure could be checked. Corpus
arguments fix the input independently of the model and make the comparison
against a manual annotation possible. A second setting on
LLM-generated arguments then shows that the same pipeline runs end-to-end on
generated input.

\section{Research Questions and Contributions}
\label{sec:intro:rqs}

This thesis builds and evaluates an end-to-end pipeline that turns
natural-language debate text into a formally evaluated abstract argumentation
framework (AF) in the sense of Dung~\cite{dung1995}. An LLM proposes
typed attack candidates between arguments, a confidence filter routes them
to acceptance, rejection, or verification, and a Prolog solver computes grounded,
preferred, and stable extensions over the accepted attacks.

The work is guided by four research questions.

\begin{description}
  \item[RQ1 (Feasibility).] Can an end-to-end pipeline be built that turns
    natural-language arguments into formal Dung extensions through LLM-based
    attack extraction and a symbolic solver?
  \item[RQ2 (Typed vs.\ binary extraction).] Does \emph{typed} attack prompting,
    which forces the model to name the attack type and justify it, produce
    different or more reliable attack relations than plain \emph{binary}
    prompting, and how does this affect the resulting frameworks?
  \item[RQ3 (Extraction quality).] How well do the extracted attacks match a
    manual reference annotation in terms of precision, recall, and F1 score,
    and which error types dominate?
  \item[RQ4 (Argumentative structure).] What argumentative structures do the
    formal semantics reveal across the eight topics of the UKP corpus
    (Table~\ref{tab:evaluation:crosstopic}), in particular regarding
    multiple extensions, and when does the formal layer add
    information beyond the grounded extension?
\end{description}

Along these questions the thesis makes four contributions: a complete and
reproducible implementation of the pipeline, from the atomization of compound
sentences into single claims through typed attack extraction and confidence
filtering to a Prolog solver (RQ1), an empirical analysis of the resulting
frameworks over eight topics of the UKP Sentential Argument Mining
Corpus~\cite{stab2018} (RQ4), an ablation that compares typed against
binary attack prompting on identical argument sets and separates the effect of
the categorization from the effect of rebuttal symmetrization (RQ2), and an
evaluation of the attack detection against a validated manual
annotation of all 201 argument pairs of one topic (RQ3).

A condensed version of the pipeline and the cross-topic evaluation has been
submitted to the International Joint Conference on Learning and Reasoning
(IJCLR 2026). In addition to the material of that submission, this thesis
contributes explicit research
questions, an expanded background, the typed-versus-binary ablation, which the
submission motivates but does not measure, and the manual reference evaluation for
RQ3.

The remainder of the thesis is structured as follows.
Chapter~\ref{sec:foundations} introduces the formal background, namely Dung's
abstract argumentation frameworks and their semantics, the attack typology of
Pollock and ASPIC+, and the basics of argument mining and language models.
Chapter~\ref{sec:related} reviews related work on LLM-based argumentation
pipelines and positions the thesis against it. Chapter~\ref{sec:method}
describes the pipeline step by step, together with the design decisions behind
it. Chapter~\ref{sec:evaluation} presents the experimental setup and results,
including the cross-topic evaluation, the typed-versus-binary ablation, the
evaluation against a manually annotated reference, and a case study with an
analysis of the recurring error types. Chapter~\ref{sec:discussion} discusses what the formal layer
adds, the limitations of the study, and the main design trade-offs, before
Chapter~\ref{sec:conclusion} summarises the answers to the research questions
and outlines future work.
